Project Rapport\\\textsc{DEVS Ex Machina}

### 2015-04-24 manual/projectrapport.tex
Add benchmarks for all formalisms to report.

\chapter{Functional requirements}
\section{DEVS Formalisms}
\section{Single core}
The single core implementations follow exactly the runtime behaviour shown by the PyPDEVS simulator. We've chosen to emulate the Select() function with an integer priority value the user can set. This saves 1 virtual function call per transition and can quite easily be optimized by the compiler since it remains constant.
\subsection{DirectConnect}
%TODO Directconnect : what is rootmodel, only atomics in core.
\subsection{Dynamic Structured}
This is implemented but removing models is not yet fully tested.
%TODO Stijn
\subsection{Parallel}
\subsubsection{Optimistic}
Optimistic time synchronization requires an implementation of timewarp, and more extensive locking than would be required for conservative. The GVT algorithm (Mattern \cite{Mattern}) runs on a separate thread from the simulation kernel threads, so care must taken that they don't deadlock, but don't race either.
\subsubsection{Conservative}
Conservative time synchronization is still under development.

\chapter{Obstacles}
\section{Virtualization}
Virtualization proved a significant hurdle by exposing potential faults. Standard compliant code was found to be leaking memory by not calling container member destructors but only on a VM with Linux as Guest and Windows as Host. This was never reproducible in another setting.
\section{Threading}
G++'s std::string implementation appears to be still Copy On Write (despite 21.4.1 p6 from the C++ standard) \cite{cpp}, triggering an alert by threadsanitizer for a possible race whenever a string was supposedly copied but internally still stored as a reference to the same object. Explicitly invoking the string constructor with the char[] proved to be a safe workaround here. Note that this problem is a known issue \cite{cow}
\section{Toolchains}
To develop we used the required toolchains:\\ g++/clang++ as compiler, Eclipse+CDT as IDE, git for dvcs, jenkins as ci, cmake as build tool supplemented with some shell and python scripting.
\section{Logger}
The recommended logger (g3log), while powerful and fast, proved a dead end since it does not compile on cygwin. 
Stijn therefore wrote his own implementation. %TODO Stijn
\chapter{Performance Analysis}
Whenever we made big design/architecture decisions, we first wrote some prototyping testcode to prove the viability (e.g. Network, Tracing, Logger).
\section{First comparison}
The first performance results were obtained by running PyPDEVS's dynamic structured example against our own simulator, with a termination time of 3.6 E6. It should be noted that the GTEST overhead is not eliminated here yet. The perf \cite{perf} tool was used to obtain low level statistics with as much detail as possible. Unfortunately perf has very limited use on a virtualized OS. The parallel implementation was not yet feature complete, so although faster could not be fairly compared to PyPDEVS.
\lstinputlisting[]{pydyn.txt}
The following shows a recording of the C++ implementation of the same example.\\
\lstinputlisting[]{cpdyn.txt}
This result is still warped due to the startup/shutdown time of GTEST, but allready the C++ port is roughly 6x faster, only page faults and migrations are worse (but being worked on).

Further optimizations included but are not limited to: 
\begin{enumerate}
\item Custom memory allocators/pools (boost, ...)
\item Profiling guided recompilation
\item Detecting bottlenecks with perf and gprof.
\item ...
\end{enumerate}

>>>>

\chapter{Functional requirements}
\section{DEVS Formalisms}
\section{Single core}
The single core implementations follow exactly the runtime behaviour shown by the PyPDEVS simulator. We've chosen to emulate the Select() function with an integer priority value the user can set. This saves 1 virtual function call per transition and can quite easily be optimized by the compiler since it remains constant.
\subsection{DirectConnect}
%TODO Directconnect : what is rootmodel, only atomics in core.
\subsection{Dynamic Structured}
This is implemented but removing models is not yet fully tested.
%TODO Stijn
\subsection{Parallel}
\subsubsection{Optimistic}
Optimistic time synchronization requires an implementation of timewarp, and more extensive locking than would be required for conservative. The GVT algorithm (Mattern \cite{Mattern}) runs on a separate thread from the simulation kernel threads, so care must be taken that they don't deadlock, but don't race either.
\subsubsection{Conservative}
Conservative time synchronization is still under development.

\chapter{Obstacles}
\section{Virtualization}
Virtualization proved a significant hurdle by exposing potential faults. Standard compliant code was found to be leaking memory by not calling container member destructors but only on a VM with Linux as Guest and Windows as Host. This was never reproducible in another setting.
\section{Threading}
G++'s std::string implementation appears to be still Copy On Write (despite 21.4.1 p6 from the C++ standard) \cite{cpp}, triggering an alert by threadsanitizer for a possible race whenever a string was supposedly copied but internally still stored as a reference to the same object. Explicitly invoking the string constructor with the char[] proved to be a safe workaround here. Note that this problem is a known issue \cite{cow}
\section{Toolchains}
To develop we used the required toolchains:\\ g++/clang++ as compiler, Eclipse+CDT as IDE, git for dvcs, jenkins as ci, cmake as build tool supplemented with some shell and python scripting.
\section{Logger}
The recommended logger (g3log), while powerful and fast, proved a dead end since it does not compile on cygwin. 
Stijn therefore wrote his own implementation. %TODO Stijn
\chapter{Performance Analysis}
Whenever we made big design/architecture decisions, we first wrote some prototyping testcode to prove the viability (e.g. Network, Tracing, Logger).
\section{First comparison}
We use perf to get statistically sound benchmarks.
\subsection{Classic}
An endtime of 3.6e6 was set, and both examples run against each other.
\lstinputlisting[]{pycla.txt}
The following shows a recording of the C++ implementation of the same example.\\
\lstinputlisting[]{cpcla.txt}
\subsubsection{Conclusion}
The C++ port is approximately 12.906 times faster.

\subsection{Dynamic structured}
An end-time of 3.6e6 was set, and both examples run against each other.
\lstinputlisting[]{pydyn.txt}
The following shows a recording of the C++ implementation of the same example.\\
\lstinputlisting[]{cpdyn.txt}
\subsubsection{Conclusion}
This result is still warped due to the startup/shutdown time of GTEST, but allready the C++ port is roughly 7.05 times faster, only page faults and migrations are worse (but being worked on).
\subsection{Parallel}
With GVT not yet active, an endtime of 3.6e6 was used (as before).\lstinputlisting[]{pypar.txt}\
The following shows a recording of the C++ implementation of the same example.\\
\lstinputlisting[]{cppar.txt}
\subsubsection{Conclusion}
The C++ port is approximately 19.6 times faster.

\section{Further work.}
Further optimizations included but are not limited to: 
\begin{enumerate}
\item Custom memory allocators/pools (boost, ...)
\item Profiling guided recompilation
\item Detecting bottlenecks with perf and gprof.
\item ...
\end{enumerate}
